Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2015-11-06 contents/implementation.tex
Updates from ShareLaTeX
updates on describing steps, shortened a lot

\noindent\textbf{Step 4:} Utilise the dense correspondences acquired by the optical flow in order to automatically generate dense and sparse (on the outline) landmark annotations and build a dense \cite{ramnath2008increasing, Amberg2009, anderson2014using} and patch-based \cite{Tzimiropoulos2014} AAM, respectively.

@@

%In practice, we assume that all the training images correspond to the same object category and contain a set of inconsistent points and curve line annotations. 
The curve annotations for all training images are densely sampled to yield a set of landmarks per image, with this set being different for every training image. To train the SVM, these landmarks are assigned as belonging to the class with label `+1', whereas randomly sampled points around them are assigned as belonging to the class with label `-1'. Since the class `+1' has far less points than the class `-1', landmarks are assigned considerably larger weights so that $N_p \times W_p=N_n \times W_n$ where $N_p, N_n$ are number of points of the class `+1' and `-1' respectively and $W_p, W_n$ are their corresponding weights.

SVMs with RBF kernel function map any number of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on 2D space represents the actual shape of the object. Note that the decision function for SVMs can be mathematically expressed as:
\begin{equation} \label{eq:decisionfunc}
    d(\bm{x})=\sum_i\alpha_i \, k(\bm{x}_i^*,\bm{x})
\end{equation}
where $\bm{x}_i^*$ are the so-called support vectors and \mbox{$k(\bm{x}, \bm{y}) = \exp\left(-\frac{\|\bm{x} -\bm{y}\|^2}{2 \sigma^2}\right)$} is the RBF kernel. The decision function $d(\bm{x})$ can be defined for every pixel $\bm{x}$ of the corresponding object image, therefore we interpret it as an image and we call it \emph{SVS image}.

Figure \ref{fig:build_svs} shows an exemplar shape representation of the human ear using SVS. As it can be observed, the final result does not drastically depend on the original number of annotated landmarks.

@@

We extend the SVS representation to support also the case where multiple curves with different labels are annotated. This is useful when annotating different structures of the same object, such as the curve of the left eyebrow and the curve of the right eyebrow on faces. Even though not absolutely necessary in our framework, it can provide further guidance on the estimation of dense shape correspondences for various object classes. In more detail, we create a multi-channel SVS image $\bm{d}(\bm{x})=[d_1(\bm{x}) \cdots d_i(\bm{x}) \cdots d_{N_c}(\bm{x})]$, where $d_i(\bm{x})$ is the SVS image that corresponds to the curve annotation of the $i$-th structure and $N_c$ is the total number of structures (a single curve annotation is the special case where $N_c=1$). Note that we do not necessarily require that all structures are annotated in all the object images: in the case that a structure is not annotated, the corresponding channel of the SVS image simply takes a zero value for all pixels. The shape flow estimation can deal with such missing information thanks to the spatial regularisation and the low-rank constraint that it adopts, c.f.~Step~\ref{sec:step3}.

@@

We define the problem of shape flow as the joint estimation of optical flow fields between a reference SVS image and every SVS image of the training dataset, which yields dense correspondences across SVS images. This also defines for every training SVS image a warping function that registers it with the reference SVS image. To establish the dense correspondences robustly, we are inspired by the idea of subspace constraints in the estimation of multiframe optical flow \cite{Garg:2013hu}.

@@

The next section describes the methodology of constructing the correspondence basis and estimating the shape flow robustly.

@@

\begin{figure}[t!]
    \centering
    \newcommand{\flowh}{0.25\columnwidth}
    \includegraphics[height=\flowh]{resources/Fig_Flows/1}
    \includegraphics[height=\flowh]{resources/Fig_Flows/2}
    \includegraphics[height=\flowh]{resources/Fig_Flows/3}
    \includegraphics[height=\flowh]{resources/Fig_Flows/4}
    \includegraphics[height=\flowh]{resources/Fig_Flows/5}
    \includegraphics[height=\flowh]{resources/Fig_Flows/6}
    \caption{Exemplar deformation fields for the left forearm, obtained using the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:deformationfield}
\end{figure}

@@

\subsubsection*{Non-rigid ICP and Correspondence Basis Creation} \label{sec:trabasis}

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply NICP \cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these  correspondence vectors and keeping only the first $R$ principal components.

%To make this correspondence vectors estimation dense (i.e.~for every pixel of the reference SVS image), we apply Thin-Plate Splines (TPS)\cite{Bookstein1989} interpolation. Finally, the correspondence basis is found by applying PCA on the result of TPS interpolation and keeping only the first $R$ principal components.

%
%
%Since optical flow pixel-wise frame registration, the trajectory basis build should also match the dimension, which is dense on shapes with trajectory length depends on number of training data. With correspondences produced after applying NICP, Thin-Plate Spline (TPS)\cite{Bookstein1989} transformations are utilized to warp every shape planes to template plane thereby generates an implicit dense shape registering. For all dense transformations $\bm{u}_n(\bm{x}), n \in \{1,...,F\}$, where $F$ is number of data and $\bm{x}$ is vector of pixels, Principle Component Analysis (PCA) is performed on trajectory to obtain low rank trajectory basis:
%\begin{equation}
%    \begin{bmatrix}
%        \bm{u_1}(\bm{x}) \\
%        \vdots \\
%        \bm{u_F}(\bm{x})
%    \end{bmatrix}
%    =
%    \begin{bmatrix}
%        \bm{q_1}(1) & \cdots & \bm{q_R}(1) \\
%        \vdots      & \ddots & \vdots  \\
%        \bm{q_1}(F) & \cdots & \bm{q_R}(F)
%    \end{bmatrix}
%    \times
%    \begin{bmatrix}
%        \bm{v_1}(x) \\
%        \vdots \\
%        \bm{v_R}(x)
%    \end{bmatrix}
%\end{equation}
%where $\bm{q_i}(n)$ are low rank components with $R \ll 2F$ and $\bm{v_i}(x)$ weighted each component with dependencies on $x$. Simpler expression shown below:
%\begin{equation}
%    \bm{u}_n(\bm{x})=\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(\bm{x})+\bm{\varepsilon_n}(\bm{x})
%\end{equation}
%Although optical flow on shapes made under the assumption that objects motion are smooth, the low rank constrain on trajectory basis recovered the hypothesis.

% \begin{figure}[t!]
%         \centering
%         \begin{tabular}{cccccccc}
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_0}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_1}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_2}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_3}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_4}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_5}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/image_6}
%         &
%         \includegraphics[trim=200 150 200 50,clip,width=.1\columnwidth]{resources/Fig_Variance/variances}
%         \end{tabular}
%     \caption{Variances}
%     \label{fig:variances}
% \end{figure}

\subsubsection*{Variational Shape Flow Estimation}%{Multi-image Subspace Flow}
\label{varshapeflow}

@@

We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu}. In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with the shape flow result of the TPS interpolation described in Sec.~\ref{sec:trabasis}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

Figure~\ref{fig:deformationfield} shows some examples of deformation fields derived from the estimated shape flow computed by the aforementioned method. These results correspond to exemplar training shapes in the case of an arm dataset.

>>>>

\noindent\textbf{Step 4:} Utilise the dense correspondences acquired by the optical flow in order to automatically generate either dense or sparse (on the outline) landmark annotations, depending on the object class type. Then, build either a dense \cite{ramnath2008increasing, Amberg2009, anderson2014using} or a patch-based \cite{Tzimiropoulos2014} AAM, respectively.

@@

%In practice, we assume that all the training images correspond to the same object category and contain a set of inconsistent points and curve line annotations.
The curve annotations for all training images are densely sampled to yield a set of landmarks per image, with this set being different for every training image. To train the SVM, these landmarks are assigned as belonging to the `positive' class, whereas randomly sampled points around them are assigned as belonging to the `negative' class. Since the positive class has far less points than the negative class, landmarks are assigned considerably larger weights so that $N_p \times W_p=N_n \times W_n$ where $N_p, N_n$ are number of points of the positive and negative class respectively and $W_p, W_n$ are their corresponding weights.

SVMs with RBF kernel functions map any number of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on the 2D space represents the actual shape of the object.
%Note that the decision function for SVMs can be mathematically expressed as:
%\begin{equation} \label{eq:decisionfunc}
%    d(\bm{x})=\sum_i\alpha_i \, k(\bm{x}_i^*,\bm{x})
%\end{equation}
%where $\bm{x}_i^*$ are the so-called support vectors and \mbox{$k(\bm{x}, \bm{y}) = \exp(-{\|\bm{x} -\bm{y}\|^2}/{2 \sigma^2})$} is the RBF kernel.
Let $d(\bm{x})$ be the decision function for SVMs. In our formulation, $d(\bm{x})$ can be defined for every pixel $\bm{x}$ of the corresponding object image, therefore we interpret it as an image and we call it \emph{SVS image}.

%Figure~\ref{fig:archi} includes some exemplar SVS representations of arm shapes (Step 1 box)

%Figure \ref{fig:build_svs} shows an exemplar shape representation of the human ear using SVS. As it can be observed, the final result does not drastically depend on the original number of annotated landmarks.

@@

We extend the SVS representation to support also the case where multiple curves with different labels are annotated. This is useful when annotating different structures of the same object, such as the left and right eyebrows on faces. Even though not absolutely necessary in our framework, it can provide further guidance on the estimation of dense shape correspondences for various object classes. In more detail, we create a multi-channel SVS image $\bm{d}(\bm{x})=[d_1(\bm{x}) \cdots d_i(\bm{x}) \cdots d_{N_c}(\bm{x})]$, where $d_i(\bm{x})$ is the SVS image that corresponds to the curve annotation of the $i$-th structure and $N_c$ is the total number of structures (a single curve annotation is the special case where $N_c=1$). Note that we do not necessarily require that all structures are annotated in all the object images: in the case that a structure is not annotated, the corresponding channel of the SVS image simply takes a zero value for all pixels. The shape flow estimation can deal with such missing information thanks to the spatial regularisation and the low-rank constraint that it adopts, c.f.~Step~\ref{sec:step3}.

@@

We define the problem of shape flow as the joint estimation of optical flow fields between a reference SVS image and every SVS image of the training dataset, which yields dense correspondences across SVS images. This also defines for every training SVS image a warping function that registers it with the reference SVS image. To establish the dense correspondences robustly, we are inspired by the idea of subspace constraints in the estimation of multiframe optical flow \cite{Garg:2013hu,tomasi2012dense,snape15faceflow}.

@@

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply Nonrigid Iterative Closest Point (NICP) \cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these correspondence vectors and keeping only the first $R$ principal components.

@@

\begin{figure}[t!]
    \centering
    \newcommand{\flowh}{0.25\columnwidth}
    \includegraphics[height=\flowh]{resources/Fig_Flows/1}
    \includegraphics[height=\flowh]{resources/Fig_Flows/2}
    \includegraphics[height=\flowh]{resources/Fig_Flows/3}
    \includegraphics[height=\flowh]{resources/Fig_Flows/4}
    \includegraphics[height=\flowh]{resources/Fig_Flows/5}
    \includegraphics[height=\flowh]{resources/Fig_Flows/6}
    \caption{Exemplar deformation fields for the left arm, obtained using the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:deformationfield}
\end{figure}

@@

@@

We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu} \footnote{We used the publicly available code at \url{http://bitbucket.org/troussos/mfsf/}}. 
We slightly modify this algorithm so that, instead of initialising the coarse-to-fine and warping iterations with a zero flow, we use Thin Plate Splines (TPS) \cite{Bookstein1989} interpolation of the initial correspondence vectors described in Step~\ref{sec:step2}.
This yields a significantly better initial location of the highly-nonconvex objective function and improves the computational efficiency, since much less coarse-to-fine pyramids are needed. 
% 
Figure~\ref{fig:deformationfield} shows some examples of deformation fields derived from the estimated shape flow computed by the aforementioned method. These results correspond to exemplar training shapes in the case of an arm dataset. We observe that the shape flow estimation captures the shape and deformations of the human arm in a plausible way.

%In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with a Thin Plate Splines (TPS) \cite{Bookstein1989} interpolation of the initial correspondence vectors described in Sec.~\ref{sec:step2}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
%$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
%$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

### 2015-11-06 contents/introduction.tex
updates on describing steps, shortened a lot

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.
%
%On the other hand,  while shapes having no motion smoothness completely as all training data are independent. Low rank constrain plays important role to simulate motion smoothness assumption.

>>>>

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu,tomasi2012dense,snape15faceflow} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.
%
%On the other hand,  while shapes having no motion smoothness completely as all training data are independent. Low rank constrain plays important role to simulate motion smoothness assumption.
